---
title: "Lab1_Advanced Stats"
output: pdf_document
---
1. What is the mean and variance for the loaded dice?
A: Mean = 4.5, Var = 3.25
```{r}
propabilities <- c(0.1,0.1,0.1,0.1,0.1,0.5)
values <- c(1,2,3,4,5,6)
expect_mean <- weighted.mean(values, propabilities)
variance <- sum((propabilities*(values^2))) - expect_mean^2

expect_mean
variance
```

2. Write a function in R that "rolls" this dice; return a vector containing the rolls
3. Make a histogram of some large number of rolls. Do the rolls of the loaded die approximate a uniform distribution?
A: No, because not all the results are equally likely, for instance, rolling a 6 has a higher probability of happening than 1-5.
```{r}
iterations <- sample(values, 10000, replace=TRUE, prob=propabilities)
hist(iterations)

iterations_func <- function(num_rolls)
{
  iteration <- sample(values, num_rolls, replace=TRUE, prob=propabilities)
  return(iteration)
}
```

4. Modify the code on Slide #58 of lecture #2 so that the means vs. trial size plots are from the loaded die.  Generate these plots a few times.  How many rolls appear to be necessary to get convergence on the expected values for the mean and variance?
A: Var & Mean = 3000 rolls

```{r}
trialSize <- c(5,10,15,20,25,30,40,50,100,200,300,400,500,1000,2000,3000,4000,5000,10000,20000,30000, 100000)
means <- vector(mode = "double", length = length(trialSize))
variances <- vector(mode = "double", length = length(trialSize))

for(i in 1:length(trialSize))
{
  rolls <- vector(length = trialSize[i], mode = "double")
  rolls <- iterations_func(trialSize[i])
  means[i] <- mean(rolls);
  variances[i] <- var(rolls)
}
plot(log10(trialSize), means)
lines(log10(trialSize), rep(4.5, length(trialSize)))
plot(log10(trialSize), variances)
lines(log10(trialSize), rep(3.25, length(trialSize)))
```

