---
title: "Lab2_Advanced Stats"
output: pdf_document
---

Question 1

1. You walk into the “occasionally dishonest casino”  with prior probabilities and likelihoods set to the values in slides 21-25 of lecture #4. (that is, a “loaded” die has a 10% chance of getting a 1-5 and a 50% chance of getting a 6 but 99% of the dice are fair)  

You pick up one die and with it roll:  

2 3 2 6 3 5 6 2 6 6 2 6 6 2 3 6 6 6 5 6 6 5 6 6 6 6 6 4 6 3 3 3 6 6 5 6 6

Make a graph of the posterior probability that you have picked up a loaded die as a function of the number of times you have rolled the die.

Show your code…

  You can represent the rolls as
data<-c(2,3,2,6,3,5,6,2,6,6,2,6,6,2,3,6,6,6,5,6,6,5,6,6,6,6,6,4,6,3,3,3,6,6,5,6,6)

```{r}
prior <- c(0.01, 0.99)
prob_loaded <- c(0.1,0.1,0.1,0.1,0.1,0.5)
prob_fair <- c(1/6,1/6,1/6,1/6,1/6,1/6)
data<-c(2,3,2,6,3,5,6,2,6,6,2,6,6,2,3,6,6,6,5,6,6,5,6,6,6,6,6,4,6,3,3,3,6,6,5,6,6)

```

```{r}
posteriors_loaded <- vector(mode = "double", length = length(data))
for(i in 1:length(data))
{
  denom <- prior[1]*prob_loaded[data[i]] + prior[2]*prob_fair[data[i]];
  prior[1] = prior[1]*prob_loaded[data[i]]/denom
  prior[2] = prior[2]*prob_fair[data[i]]/denom
  posteriors_loaded[i] <- prior[1]
}
print(posteriors_loaded)
```
```{r}
indices <- 1:length(posteriors_loaded)
probabilities_loaded <- posteriors_loaded
plot(indices,probabilities_loaded)

```


2. How many times on average would you need to roll a loaded die to be 99.999%  sure that it was loaded?  
(Show your work)

A: 56.6278

```{r}
prior <- c(0.01, 0.99)
prob_loaded <- c(0.1,0.1,0.1,0.1,0.1,0.5)
prob_fair <- c(1/6,1/6,1/6,1/6,1/6,1/6)

dataoflikelihoodld <- function(value, likelihood, numruns)
{
   datad <- sample(value, numruns, prob = likelihood)
   return(datad)
}

repeatrunld <- function(numruns)
{
    totalruns <- vector();
    for(j in 1:numruns)
    {
      prior <- c(0.01, 0.99)
      values <- c(1,2,3,4,5,6)
      prob_loaded <- c(0.1,0.1,0.1,0.1,0.1,0.5)
      prob_fair <- c(1/6,1/6,1/6,1/6,1/6,1/6)
      probload <- vector();
      probfair <- vector();
      x <- vector();
      indexes = 1;
      while(prior[1] <= 0.99999)
      {
        data <- dataoflikelihoodld(values, prob_loaded, numruns = 1)
        x[indexes] <- indexes
        denom <- prior[1]*prob_loaded[data] + prior[2]*prob_fair[data];
        prior[1] = prior[1]*prob_loaded[data]/denom
        prior[2] = prior[2]*prob_fair[data]/denom
        probload[indexes] <- prior[1]
        probfair[indexes] <- prior[2]
        indexes <- indexes + 1;
      }
      totalruns[j] <- tail(x, n=1)
    }
  hist(totalruns)
  return(mean(totalruns))
}
print(repeatrunld(numruns = 10000))
```




Question 2

You are consulting for a hospital.  They have a diagnostic test for a disease with a known background prevalence of 0.1%.

The test has the following properties:
p(positive result | person has disease) = 0.91
p(negative result| person does not have disease) = 0.84

The cost of running the test one time is $1.  The test can be repeated for each patient and the results of the
test are independent of one another allowing for Bayesian updates.  The test always yields a positive or negative 
result.

The requirement of the hospital is that the test is repeated for each patient until a Bayesian posterior 
of at least 0.99999 is reached.

```{r}
# Positive = 0.91, 0.09  Negative = 0.84, 0.16

disease <- c(0.91, 0.09)
nodisease <- c(0.16, 0.84)

dataoflikelihood <- function(likelihood, numruns)
{
  data <- vector(mode="integer", length=numruns);
  for(i in 1:numruns)
  {
    if(runif(1) <= likelihood[1])
    {
      data[i] <- 1;
    }
    else
    {
      data[i] <- 2;    
    }
  }
  return(data)
}

postprobd <- 0.001
postprobnd <- 1 - postprobd
probdisease <- vector();
probnodisease <- vector();
ynvalues <- vector();
x <- vector();
indexes = 1;
while(postprobd <= 0.99999)
{
   data <- dataoflikelihood(disease, 1)
   x[indexes] <- indexes
   ynvalues[indexes] <- data;
   probdisease[indexes] <- postprobd;
   probnodisease[indexes] <- postprobnd;
   denom <- postprobd * disease[data] + postprobnd * nodisease[data];
   postprobd = postprobd * disease[data]/denom;
   postprobnd = postprobnd * nodisease[data]/denom;
   
   #plot(1:indexes,probdisease, ylim=c(0,1), xlim=c(1, indexes + 1))
   #Sys.sleep(1);
   indexes <- indexes + 1;
}

plot(x, probdisease)

```



1. Run simulations for a patient with the disease.  About how many times on average must the test be repeated to achieve the hospital’s requirements?  

A: 13.8425
```{r}
repeatrunsp <- function(numruns)
{
    totalruns <- vector();
    for(j in 1:numruns)
    {
      postprobd <- 0.001
      postprobnd <- 1 - postprobd
      probdisease <- vector();
      probnodisease <- vector();
      x <- vector();
      indexes = 1;
      while(postprobd <= 0.99999)
      {
        data <- dataoflikelihood(disease, 1)
        x[indexes] <- indexes
        probdisease[indexes] <- postprobd;
        probnodisease[indexes] <- postprobnd;
        denom <- postprobd * disease[data] + postprobnd * nodisease[data];
        postprobd = postprobd * disease[data]/denom;
        postprobnd = postprobnd * nodisease[data]/denom;
   
        #plot(1:indexes,probdisease, ylim=c(0,1), xlim=c(1, indexes + 1))
        #Sys.sleep(1);
        indexes <- indexes + 1;
      }
      totalruns[j] <- tail(x, n=1)
    }
  hist(totalruns)
  return(mean(totalruns))
}

print(repeatrunsp(numruns = 10000))
```


2. Repeat the simulations for a patient without the disease.  About how many times on average must the test be repeated to achieve the hospital’s requirements?

A: 3.7923
```{r}
#P(no disease | negative result) = p(ngative result | no disease)*p(no disease)/p(negative result)
disease <- c(0.91, 0.09)
nodisease <- c(0.16, 0.84)

repeatrunsn <- function(numruns)
{
    totalruns <- vector();
    for(j in 1:numruns)
    {
      postprobd <- 0.001
      postprobnd <- 1 - postprobd
      probdisease <- vector();
      probnodisease <- vector();
      x <- vector();
      indexes = 1;
      while(postprobnd <= 0.99999)
      {
        data <- dataoflikelihood(nodisease, 1)
        x[indexes] <- indexes
        probdisease[indexes] <- postprobd;
        probnodisease[indexes] <- postprobnd;
        denom <- postprobd * disease[data] + postprobnd * nodisease[data];
        postprobd = postprobd * disease[data]/denom;
        postprobnd = postprobnd * nodisease[data]/denom;
   
        #plot(1:indexes,probdisease, ylim=c(0,1), xlim=c(1, indexes + 1))
        #Sys.sleep(1);
        indexes <- indexes + 1;
      }
      totalruns[j] <- tail(x, n=1)
    }
  hist(totalruns)
  return(mean(totalruns))
}

print(repeatrunsn(numruns = 10000))
```


3. The hospital plans to run the test on one million patients per year.  At a cost of $1 per test, about how much
should the hospital budget to run these tests?  (That is to say, for a million patients, how many tests can 
the hospital anticipate running?)

A: $3,892,802
```{r}
patients <- vector()

testsposave <- 13.8425
testsnave <- 3.7923
prior <- 0.01

calcprice <- function(testsposave, testsnave, prior)
{
    value <- testsposave*prior+testsnave*(1-prior)
    cost <- value*1000000
    return(cost)
}

calcprice(testsposave, testsnave, prior)

```

