---
title: "Lab4_Advanced_Stats"
output: pdf_document
---

1) Read the dataset into R using commands something like...

setwd("C:\\classes\\Advanced_Stats_Spring2022\\Lab4_HW1")

myT <- read.table("nc101_scaff_dataCounts.txt",header=TRUE,row.names=1)


(2) On a log10-log10 scale, show a plot of the counts for the two samples “D2_01” and “D2_02”.

```{r}
setwd("/Users/ciaraconway/Documents/R_Files")
dataT <- read.table("nc101_scaff_dataCounts.txt", header=TRUE, row.names = 1)

dataT$D2_01_log = log10(dataT$D2_01)
dataT$D2_02_log = log10(dataT$D2_02)

head(dataT)
```

(2) Plot

A: Yes, it appears that overall, when log transformed, there seems to be a somewhat linear relationship between D2_01 and D2_02.  However, with the earlier points, there seems to be a devation from that relationship
```{r}

plot(dataT$D2_01_log, dataT$D2_02_log, col="darkmagenta")
```
(3) On a log10-log10 scale, plot the variance of all genes (across all samples) vs. the mean (across all genes) with a red line on your graph representing the identity line.  Does the mean equal the variance for these samples?   
(The commands apply(myT, 1, mean) and apply(myT, 1, var) will be of some use).  

```{r}
dataT2 <- read.table("nc101_scaff_dataCounts.txt", header=TRUE, row.names = 1)
#dataT2[,1:11] <- log10(dataT2[,1:11]+1)

dataMean <- apply(dataT2, 1, mean)
dataVar <- apply(dataT2, 1, var)

dataMean <- log10(dataMean)
dataVar <- log10(dataVar)

plot(dataMean, dataVar, col="blue")
#add identity line
lines(c(-6,6), c(-6,6), col="red")
```
(4) Consider the first gene in the spreadsheet (e.g. NC101_00003).  Make a two by two contingency table:

```{r}
col_labels <- c("Assigned to NC101_00003", "Not assigned to NC101_00003")
Seq_in_D2_01 <- c(dataT$D2_01[1], sum(dataT$D2_01) - dataT$D2_01[1])
Seq_in_D2_02 <- c(dataT$D2_02[1], sum(dataT$D2_02) - dataT$D2_02[1])

dataCon <- data.frame(col_labels, Seq_in_D2_01, Seq_in_D2_02)
dataCon
```
use the two sided fisher.test to generate a p-value for the null hypothesis that the columns and rows of the contingency table are independent.

```{r}
m <- matrix(c(dataT$D2_01[1], sum(dataT$D2_01) - dataT$D2_01[1], dataT$D2_02[1], sum(dataT$D2_02) - dataT$D2_02[1]), nrow = 2)
fisher.test(m, alternative = "two.sided")

```
(5) Now generate a p-value for all the genes in the spreadsheet from the Fisher test.  Plot out those p-values in a histogram.  Are they uniformly distributed?  Would you expect them to be?  Are the p-values more significant, less significant or what we would expect under a uniform distribution?  How does the p-value distribution change if you remove low abundance genes (with for example myT <- myT[ (myT$D2_01 + myT$D2_02 > 50),]

```{r}
pvalues <- vector()
for(i in 1:nrow(dataT))
{
    m <- matrix(c(dataT$D2_01[i], sum(dataT$D2_01) - dataT$D2_01[i], dataT$D2_02[i], sum(dataT$D2_02) -  dataT$D2_02[i]), nrow = 2)
    ft <- fisher.test(m, alternative = "two.sided")
    pvalues[i] = ft$p.value
}

hist(pvalues, col = 'skyblue3', breaks = 50)
```



```{r}
dataT3 <- dataT[(dataT$D2_01 + dataT$D2_02 > 50), ]

pvaluesR <- vector()
for(i in 1:nrow(dataT3))
{
    m <- matrix(c(dataT3$D2_01[i], sum(dataT3$D2_01) - dataT3$D2_01[i], dataT3$D2_02[i], sum(dataT3$D2_02) -  dataT$D2_02[i]), nrow = 2)
    ft <- fisher.test(m, alternative = "two.sided")
    pvaluesR[i] = ft$p.value
}

hist(pvalues, col = 'skyblue3', breaks = 50)
```
(6) Add 1 to every value in the table ( with something like myT = myT + 1 ).  This is called adding a pseudo-count.  Now consider the first gene (NC101_00003 ) again.  From the first experiment, calculate 
	
expected frequency = p = 
		(# Assigned to NC101_00003 in D2_01)/total # of sequences in D2_01)

Now use poisson.test to assign a p-value for the null hypothesis that value of p derived from D2_01 could have produced the number of reads observed for this gene in D2_02 

```{r}
dataT4 <- dataT + 1
exp_freq <- dataT4$D2_01[1]/sum(dataT4$D2_01)
poisson.test(dataT4$D2_01[1], sum(dataT4$D2_01), r = exp_freq)
poisson.test(dataT4$D2_02[1], sum(dataT4$D2_02), r = exp_freq)
```

(7) Repeat the calculation in (6) for every gene in the spreadsheet.  Graph these p-values against the p-values produced in (5) on a log10-log10 plot.  How well do they agree? 

Answer: They somewhat agree, I see similar linearity, but they are not identical.

```{r}
### Comparing D2_01 to D2_02
pvaluesPN <- vector()

for(i in 1:nrow(dataT4))
{
    exp_freq3 <- dataT4$D2_01[i]/sum(dataT4$D2_01)
    pt <- poisson.test(dataT4$D2_02[i], sum(dataT4$D2_02), r=exp_freq3)
    pvaluesPN[i] <- pt$p.value
}
```

```{r}
pvaluesL <- log10(pvalues)
pvaluesPNL <- log10(pvaluesPN)
plot(pvaluesPNL, pvaluesL, col="steelblue")

#add identity line
#lines(c(-400,400), c(-400,400), col="red")
```

