---
title: "Lab5_Advanced_Stats"
output: pdf_document
---
(1A): Download the data from here examining the relationship between the number of cell divisions and cancer risk: https://fodorclasses.github.io/classes/stats2020/cancerRisk.txt
On a log10-log10 scale graph Lifetime_cancer_risk (on the y-axis) vs. CumulativeCellDivisions (on the x-axis).  (This reproduces Fig. 1 from the paper).

```{r}
cancerFile <- "cancerRisk.txt"
cancerData <- read.table(cancerFile, header = TRUE)

plot(log10(cancerData$CumulativeCellDivisions), log10(cancerData$Lifetime_cancer_risk), pch=21, col= "red4", bg= "seagreen3")

```

(1B):   Using the lm function, fit a linear model with Lifetime_cancer_risk as the Y variable and CumulativeCellDivisions as the x-data.  Add the regression line to the plot using the function abline(myLm)  (where myLm is the linear model you created).

```{r}
linearMod <- lm(CumulativeCellDivisions~Lifetime_cancer_risk, data = cancerData)
linearModlog <- lm(log10(CumulativeCellDivisions)~log10(Lifetime_cancer_risk), data = cancerData)

linearModlog$coefficients
plot(log10(cancerData$CumulativeCellDivisions), log10(cancerData$Lifetime_cancer_risk), pch=21, col= "red4", bg= "seagreen3")
abline(a = linearModlog$coefficients[1], b = linearModlog$coefficients[2])

```
(1C):  What is the p-value for the null hypothesis that the slope of the regression between these two variables is zero?  What is the r-squared value of the model?

Answer = p-value: 0.002028, Adjusted R-squared:  0.2592 

```{r}
print(linearMod)
summary(linearMod)
```

(1D): Are the assumptions of constant variance and normal distribution of the residues reasonable for this model?  Justify your answer.
Answer = by looking at the Residuals vs. Fitted plot you cans see that there is relatively constant variance.  by looking at the Normal Q-Q plot you can see that the data is someone normally distributed

```{r}
plot(linearModlog)
```

(2) Consider the case-control file for the colorectal adenomas data set that is here:
	http://afodor.github.io/classes/stats2015/caseControlData.txt
    A separate file gives obesity (BMI) data for these same subjects:
	http://afodor.github.io/classes/stats2015/BMI_Data.txt

```{r}
bmi <- read.table("BMI_Data.txt", sep = "\t", header = TRUE)
case <- read.table("caseControlData.txt", sep = "\t", header = TRUE)

### Python Merge ###
new_data <- read.csv("bmi_case_merge.csv", header = TRUE)

```

For each OTU in the spreadsheet, generate a p-value from linear regression comparing BMI to the   relative abundance of each OTU.    Graph out all the p-values.  Do they appear uniformly distributed? Does the microbial community appear to be influencing body weight in this cohort?  Are any of these associations significant at a 10% false discovery rate?
Hints:  To lookup the ids in the BMI table, you will need to some processing on the “sample” column in the caseControl file.  The following code will convert the a sampleID so that it will match the BMI file.
	# remove case and control
key <- sub("case", "", sampleID)
	key <- sub("control", "", key)
	
	# remove extraneous information from the suffix
	key <- strsplit( key, "_")[[1]][1]

Also, to get the p-value out of the linear model try:
	anova( myLm)$"Pr(>F)"[1]
We’ll see why that work shortly in future lectures.

```{r}

lmvector <- vector()
for(i in 2:ncol(new_data)-1)
{
  #print(new_data[i])
  lm_entry <- lm(new_data$bmi~new_data[,i], data = new_data)
  #pval <- aov(lm_entry)$"Pr(>F)"[1]
  pvall <- anova(lm_entry)$"Pr(>F)"[1]
  lmvector[i-1] <- pvall
}
print(lmvector)
```

Histogram - PVal
```{r}
#lmvector <- p.adjust(lmvector)
hist(lmvector, breaks = 20)
```

P Val Adjustments
```{r}
pval_adj <- p.adjust(lmvector, method = "BH")
#pval_adj_10 <- p.adjust(lmvector, method = "")
hist(pval_adj)
```
10% False Discovery

```{r}
pval_adj_10 <- pval_adj[pval_adj<= 0.10]
hist(pval_adj_10)
```
Do they appear uniformly distributed? Does the microbial community appear to be influencing body weight in this cohort?  Are any of these associations significant at a 10% false discovery rate?

Answer = The un-adjusted values seem to be in a somewhat of a uniform distribution. Some seem to hold significance, yes.  No they are not.
